A generalized definition of multidimensional item response theory parameters

### 2024-03-28 output/paper_draft/paper_M2PLM_parameters.Rmd
Applies approved changes in "Model Formulation"
Changes approved by me and @mluzardov in the Google Docs version

In the non-cognitive domain, the use of so-called negatively-keyed items
---items that tap the negative pole of a certain trait, and thus
usually reverse-scored--- is also widespread.<!--# Reference and example? -->
It is common to find instruments that combine
positively- and negatively-keyed items.
Moreover, a complex item may have discrimination parameters with opposite signs
for different latent dimensions.<!--# Examples? -->
Take the multidimensional version of the nominal categories model
[@thissen_nominal_2010] as an example,
and consider two response categories tapping a different construct each:
A positive change in the dimension tapped by one category will inevitably
lead to a decrease in the probability of response to the other category.
One can easily note that a monotonically decreasing probability for a certain
dimension is given by a non-positive discrimination parameter
(and an unchanging probability by a null discrimination parameter),<!--# Actually, this also applies to the cognitive domain, which may not be very well explained. -->
contrary to the constraints imposed by @reckase_difficulty_1985 [p. 411].
However, the multidimensional item parameters can also be defined for
items with monotonically non-increasing probabilities, as we will see.

@@

Consider a $`r N_DIMS`$-dimensional latent space $`r LATENT_SPACE`$
in basis $`r LS_BASIS_EQ`$.
A respondent is represented by vector $`r TRAIT_VEC_IN_LS`$,
which is assumed to be a multivariate normal random variable:

\begin{equation}
  `r TRAIT_MV_DEF`,
\end{equation}

where $`r MEAN_VECTOR`$ is an $n$-element vector of means and
$`r COV_MATRIX` = `r SD_MATRIX` `r CORR_MATRIX` `r SD_MATRIX`$
an $n$-dimensional positive definite covariance matrix,
with $`r CORR_MATRIX`$ the corresponding correlation matrix
(also positive definite) and $`r SD_MATRIX`$ a scaling diagonal matrix
with $`r VAR_COV_EQ`$ (i.e., the variances).

According to the `r MODEL_ACRONYM` model, the probability of
a positive response to item $`r ITEM_INDEX`$ is

@@

Our aim is to find the point of maximum slope of $`r IRF_ABBR`$ in
$`r LATENT_SPACE`$.
Because the slope is dependent on the direction considered,
we need to reparameterize Equation \@ref(eq:M2PL-formulation)
in polar coordinates, as in @reckase_difficulty_1985.

>>>>

In the non-cognitive domain, the use of so-called negatively-keyed items
---items that tap the negative pole of a certain trait, and thus
usually reverse scored--- is also widespread.<!--# Reference and example? -->
It is common to find instruments that combine
positively- and negatively-keyed items.
Moreover, a complex item may have discrimination parameters with opposite signs
for different latent dimensions.<!--# Examples? -->
Take the multidimensional version of the nominal categories model
[@thissen_nominal_2010] as an example,
and consider two response categories tapping a different construct each:
A positive change in the dimension tapped by one category will inevitably
lead to a decrease in the probability of response to the other category.
One can easily note that a monotonically decreasing probability for a certain
dimension is given by a non-positive discrimination parameter
(and an unchanging probability by a null discrimination parameter),<!--# Actually, this also applies to the cognitive domain, which may not be very well explained. -->
contrary to the constraints imposed by @reckase_difficulty_1985 [p. 411].
However, the multidimensional item parameters can also be defined for
items with monotonically non-increasing probabilities, as we will see.

@@

Consider a $`r N_DIMS`$-dimensional latent space $`r LATENT_SPACE`$
in basis $`r LS_BASIS_EQ`$, being $`r LS_BASIS`$ not necessarily orthonormal,
where a respondent is represented by vector $`r TRAIT_VEC_IN_LS`$.
According to the `r MODEL_ACRONYM` model, the probability of
a positive response to item $`r ITEM_INDEX`$ is

@@

Our aim is to find in the first place
the point of maximum slope of $`r IRF_ABBR`$ in $`r LATENT_SPACE`$.
Because the slope is dependent on the direction considered,
we need to reparameterize Equation \@ref(eq:M2PL-formulation)
in polar coordinates, as in @reckase_difficulty_1985.
